---
title: "PracticalMachineLearning"
author: "Necula Calin"
date: "02/22/2015"
output: html_document
---


### Getting and Cleaning Data + Exploratory Analysis
```{r} 
setwd("/home/Calin/Desktop/coursera/PracticalMachineLearning/")
# download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","train.csv", method="curl")
# download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv","test.csv", method="curl")
library(rpart)
# Read the data 
train <- read.csv("train.csv", na.strings=c("NA",""), header=TRUE)
test <- read.csv("test.csv", na.strings=c("NA",""), header=TRUE)
```

| Checking the NA distribution
```{r}
summary(colSums(is.na(train)))
summary(colSums(is.na(test)))
```

| Most variables are NA on all lines so we remove them
| Removing the NA lines

```{r}
train<-train[,colSums(is.na(train)) == 0]
test<-test[,colSums(is.na(test)) == 0]
```
First variables have no use in prediction.
Removing the "header" rows
```{r}
train<-train[,-c(1:7)]
test<-test[,-c(1:7)]
```

Last collumn is for classe and problem ID I remove the problem ID column
```{r}
test<-test[,1:(ncol(test)-1)]
```

Split the data in a train subset and a validation subset
```{r}
library(caret)
inTrain <- createDataPartition(y=train$classe, p=0.75, list=FALSE)
Train <- train[inTrain, ] 
validation <- train[-inTrain, ]
```

###Using Decision Tree
```{r}
library(rpart)
mod1 <- rpart(classe ~ ., data=Train, method="class")
pred1 <- predict(mod1, validation, type = "class")
confusionMatrix(pred1, validation$classe)
```

###Using Random Forest from caret
```{r}
library(randomForest)
mod2<-randomForest(classe ~. , data=Train, method="class")
pred2 <- predict(mod2, validation, type = "class")
confusionMatrix(pred2, validation$classe)
```
### Conclusion
The accuracy of the random forest model is `99.5%` with a confidence interval of `99.3%` to `99.7%`, it is better then the Decision Trees model with `75.9%`.

| I choose the randomForest model because it has better accuracy.

###Getting the submission results
```{r}
submission <- predict(mod2, test, type="class")
```
### Writing the files
```{r}
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

pml_write_files(submission)
```